On finiteness of structures produced deterministically by co-transcriptional folding

### 2018-11-28 ttt.tex
update 11/28 in Tokyo

            \node[below] at (180:1) {$w[i-2]$};
            \node[above] at (0:0) {$w[i-1]$};
            \node[right] at (-120:1) {$w[i]$};

@@

    \end{tabular}
    \caption{Case of $S[i-2..i] = tbt$}
    \label{TTT_tunnelC_enter_usingTunnel}
\end{figure}

\subsubsection{Case of $S[i-2..i] = bbt$}
Let us consider number of consumed by $w[i-2]$ (Fig.\ref{TTT_tunnelC_enter_usingBond}).

\begin{itemize}
\item{$w[i-2]$ consumes one binding}\\
  In this situation, $w[i-2]$ supplies one active binding whereas $w[i]$ consumes this binding. In addition, $w[i-1]$ has to bound to one of $n_1, n_2, n_3$.
  Thus, in this cases consume some binding capabilities.

\item{$w[i-2]$ consumes two bindings}\\
  In this case, $w[i-2]$ already consumes two binding. $w[i-1]$ has to be bound. $w[i]$ supplies at most two bindings. Thus, in this cases consume some binding capabilities.
\end{itemize}

@@

          \node[below left] at (180:1) {$w[i-2]$};
          \node[above] at (0:0) {$w[i-1]$};
          \node[below] at (-120:1) {$w[i]$};

@@

          \end{scope}
        \end{tikzpicture}
    \caption{Case of $S[i-2..i] = bbt$}
    \label{TTT_tunnelC_enter_usingBond}
  \end{center}
\end{figure}

\subsubsection{$\alpha = 2$}

By Tunnel Troll Theorem, any tunnel sections which represented in $bbt^+$ or $bt^+bt^+$ consume binding capabilities. If the sequence $S$ is free from any subsequence of the form $bt^+bt^+$, then it can factorize as $S = u_1 u_2 u_3 \cdots$ for some $u_1 , u_2 , u_3 , \cdots \in \{b\} \cup bbt^+$. Assume the length of $\sigma$ is $n$, seed supplies at most $2n$ binding capabilities. Therefore formula \ref{TTT_only_bond} hold.

@@

Let us consider the case of $v_k$ is $o$. See Fig.\ref{TTT_case_of_o}. $w[i-1]$ consumes some binding capabilities because $v_{i-1}$ is $b$. If the number of $w[i-1]$'s bindings is one binding, then $w[i+1]$ has to be bound except $n_1$ or $n_2$ so that $w[i+1]$ must consumes two bindings except the case of $n_1$ and $n_2$ are occupied and $w[i]$ consumes at least one binding. If $n_1$ and $n_2$ are occupied, then $w[i-1]$'s bindings are inactive that is $w[i-1]$ consumes two binding capabilities. Therefore, this case consumes binding capabilities. If $w[i-1]$ dose Not have any bindings, then $w[i-1]$ already consumes two bindings. In addition, $w[i]$ and $w[i+1]$ consume at least one binding. Therefore this case consumes binding capabilities. Thus, the formula \ref{TTT_only_a} hold and according to the formula \ref{TTT_only_bond} and the formula \ref{TTT_only_a}, the formula \ref{TTT_structure} is hold.

@@

\subsubsection{$\alpha = 3$}

\begin{theorem}[$\delta = 1, \alpha = 3$]
Let $\Xi$ be a unary oritatami system of $\delta = 1, \alpha = 3$. It can yield only a finite structure whose size is $\mathcal{O}(n)$.
\end{theorem}

@@

\begin{proof}[proof of lemma]
Assume the transcript is stabilized until $w[i-1]$. One of neighbors of $p$ is not $w[i-1]$ where this bead regards $n_1$. If $w[i-1]$ include neighbors of $p$ and $w[i]$ is stabilized at another point of $p$ with one bond. Then, any neighbors do not have bond without $w[i-1]$. Neighbors of $n_1$ have to be occupied at least five according to lemma \ref{TTT_neighbor_lemma} and two of them include neighbors of $p$ where each of them regards $n_2, n_3$. In the same way, five neighbors of $n_2$ and $n_3$ are occupied and each of one of them includes neighbors of $p$ where they regard $n_4, n_5$. one of $n_5$'s neighbors includes neighbors of $p$ where it regards $n_6$. Then, any neighbors of $p$ are occupied. That is if some neighbors of $p$ are free, then there exists a bead which has bonds in neighbors.
\end{proof}

\begin{proof}
Let us show that $\#bc(C_{i-1}) > \#bc(C_i)$, that is when $w[i]$ is stabilized, $w[i]$ uses at least two hands.
Let us assume $w[i]$ is able to be stabilized with using one hand. Fig.\ref{TTT_a3_w} exhibits all the three kinds of possibility of stabilized $w[i]$. Then, $w[i]$ can be also stabilized at $n_3$.

@@

\item[-] Otherwise\\
Because of  $S[i] = b$, at least one of $n_1$ and $n_2$ have to be free. Let us regard that $n_1$ is free. Neighbors of $n_1$ have to be occupied and at least two neighbors of $n_{-1}$ have to be free for $n_1$ and $w[i]$. According to lemma\ref{TTT_neighbor_lemma}, $n_{-1}$ have some hand. Therefore $w[i]$ can be also stabilized $n_1$ that is this situation is non-deterministic. Thus, one of $n_3$ and $n_4$ has to be free.
\end{itemize}

@@

\subsubsection{$\alpha = 4$}
\begin{theorem}[$\delta = 1, \alpha = 4$]
Let $\Xi$ be a unary oritatami system of $\delta = 1, \alpha = 4$. It can yield only a finite structure whose size is $\mathcal{O}(n^2)$.
\end{theorem}

@@

      \draw[dashed] (180:2) -- (0:2);
      \draw[->, blue] (-120:0.9) -- (-120:0.1);

@@

	\node at (0:2.5) {$n+1$};
    \end{tikzpicture}
    \caption{the first bead out of $\hexagon_{w[-n+1]}^n$}
    \label{TTT_a4_first}
\end{figure}

>>>>

            \node[below] at (180:1) {$w[h-1]$};
            \node[above] at (0:0) {$w[h]$};
            \node[right] at (-120:1) {$w[h+1]$};

@@

    \end{tabular}
    \caption{Case of $S[h-1..h+1] = tbt$}
    \label{TTT_tunnelC_enter_usingTunnel}
\end{figure}

\subsubsection{Case of $S[h-1..h+1] = bbt$}
Let us consider number of consumed by $w[h-1]$ (Fig.\ref{TTT_tunnelC_enter_usingBond}).

\begin{itemize}
\item{$w[h-1]$ consumes one binding}\\
  In this situation, $w[h-1]$ supplies one active binding whereas $w[h+1]$ consumes this binding. In addition, $w[h]$ has to bound to one of $n_1, n_2, n_3$.
  Thus, in this cases consume some binding capabilities.

\item{$w[h-1]$ consumes two bindings}\\
  In this case, $w[h-1]$ already consumes two binding. $w[h]$ has to be bound. $w[h+1]$ supplies at most two bindings. Thus, in this cases consume some binding capabilities.
\end{itemize}

@@

          \node[below left] at (180:1) {$w[h-1]$};
          \node[above] at (0:0) {$w[h]$};
          \node[below] at (-120:1) {$w[h+1]$};

@@

          \end{scope}
        \end{tikzpicture}
    \caption{Case of $S[h-1..h+1] = bbt$}
    \label{TTT_tunnelC_enter_usingBond}
  \end{center}
\end{figure}

\end{proof}

\subsection{On structures provided by a unary and $\delta = 1$ oritatami system}
\begin{theorem}[$\delta= 1, \alpha=2$]
Let $\Xi$ be a unary oritatami system of $\delta = 1, \alpha = 2$. It can yield infinite structures but they are only zig-zag conformation.
\end{theorem}

\begin{proof}
By Tunnel Troll Theorem, any tunnel sections which represented in $bbt^+$ or $bt^+bt^+$ consume binding capabilities. If the sequence $S$ is free from any subsequence of the form $bt^+bt^+$, then it can factorize as $S = u_1 u_2 u_3 \cdots$ for some $u_1 , u_2 , u_3 , \cdots \in \{b\} \cup bbt^+$. Assume the length of $\sigma$ is $n$, seed supplies at most $2n$ binding capabilities. Therefore formula \ref{TTT_only_bond} hold.

@@

Let us consider the case of $v_k$ is $o$. See Fig.\ref{TTT_case_of_o}. $w[i-1]$ consumes some binding capabilities because $v_{i-1}$ is $b$. If the number of $w[i-1]$'s bindings is one binding, then $w[i+1]$ has to be bound except $n_1$ or $n_2$ so that $w[i+1]$ must consumes two bindings except the case of $n_1$ and $n_2$ are occupied and $w[i]$ consumes at least one binding. If $n_1$ and $n_2$ are occupied, then $w[i-1]$'s bindings are inactive, that is, $w[i-1]$ consumes two binding capabilities. Therefore, this case consumes binding capabilities. If $w[i-1]$ dose Not have any bindings, then $w[i-1]$ already consumes two bindings. In addition, $w[i]$ and $w[i+1]$ consume at least one binding. Therefore this case consumes binding capabilities. Thus, the formula \ref{TTT_only_a} hold and according to the formula \ref{TTT_only_bond} and the formula \ref{TTT_only_a}, the formula \ref{TTT_structure} is hold. Thus, in this case, oritatami system can yield infinite structures but they are only zig-zag conformation.

\end{proof}

@@

\begin{theorem}[$\delta = 1, \alpha = 3$]
Let $\Xi$ be a unary oritatami system of $\delta = 1, \alpha = 3$. It can yield only finite structures whose size is $\mathcal{O}(n)$.
\end{theorem}

@@

\begin{proof}[proof of lemma]
Assume the transcript is stabilized until $w[i-1]$. One of neighbors of $p$ is not $w[i-1]$ where this bead regards $n_1$. If $w[i-1]$ include neighbors of $p$ and $w[i]$ is stabilized at another point of $p$ with one bond. Then, any neighbors do not have bond without $w[i-1]$. Neighbors of $n_1$ have to be occupied at least five according to lemma \ref{TTT_neighbor_lemma} and two of them include neighbors of $p$ where each of them regards $n_2, n_3$. In the same way, five neighbors of $n_2$ and $n_3$ are occupied and each of one of them includes neighbors of $p$ where they regard $n_4, n_5$. one of $n_5$'s neighbors includes neighbors of $p$ where it regards $n_6$. Then, any neighbors of $p$ are occupied. That is, if some neighbors of $p$ are free, then there exists a bead which has bonds in neighbors.
\end{proof}

\begin{proof}
Let us show that $\#bc(C_{i-1}) > \#bc(C_i)$, that is, when $w[i]$ is stabilized, $w[i]$ uses at least two hands.
Let us assume $w[i]$ is able to be stabilized with using one hand. Fig.\ref{TTT_a3_w} exhibits all the three kinds of possibility of stabilized $w[i]$. Then, $w[i]$ can be also stabilized at $n_3$.

@@

\item[-] Otherwise\\
Because of  $S[i] = b$, at least one of $n_1$ and $n_2$ have to be free. Let us regard that $n_1$ is free. Neighbors of $n_1$ have to be occupied and at least two neighbors of $n_{-1}$ have to be free for $n_1$ and $w[i]$. According to lemma\ref{TTT_neighbor_lemma}, $n_{-1}$ have some hand. Therefore $w[i]$ can be also stabilized $n_1$, that is, this situation is non-deterministic. Thus, one of $n_3$ and $n_4$ has to be free.
\end{itemize}

@@

\begin{theorem}[$\delta = 1, \alpha = 4$]
Let $\Xi$ be a unary oritatami system of $\delta = 1, \alpha = 4$. It can yield only finite structures whose size is $\mathcal{O}(n^2)$.
\end{theorem}

@@

      \draw[dashed] (180:2) -- (0:2);
      \draw[dashed, shift=(-60:1)] (180:2.5) -- (0:1.5);
      \draw[->, blue] (-120:0.9) -- (-120:0.1);

@@

	\node at (0:2.5) {$n+1$};
	\node[shift=(-60:1)] at (0:2) {$n$};
    \end{tikzpicture}
    \caption{the first bead out of $\hexagon_{w[-n+1]}^n$}
    \label{TTT_a4_first}
\end{figure}
